\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2019-06-21 draft.tex
apply some suggestions from Tomas

\subsection*{Related Works}
Reducer commutativity problem can be reduced to a program equivalence problem. One creates another program $R'$ that first non-deterministically swaps two consecutive input values and then executes the code of $R$. If $R'$ and $R$ are equivalent, using the fact that all permutations of a list can be obtained by swapping consecutive list elements finitely many times\footnote{Here is an example to produce [3;2;5;1;4] from [1;2;3;4;5] by swapping consecutive elements: [\textbf{1};\textbf{2};3;4;5] $\rightarrow$ [2;\textbf{1};\textbf{3};4;5] $\rightarrow$ [2;3;1;\textbf{4};\textbf{5}]$\rightarrow$ [2;3;\textbf{1};\textbf{5};4]$\rightarrow$ [\textbf{2};\textbf{3};5;1;4]$\rightarrow$ [3;2;5;1;4].}, $R$ can be proved to be commutative. There is a series of research results on program equivalence checking (or closely related topics such as regression verification and translation validation), to name a few~\cite{Pnueli:1998:TV,fedyukovich2015automated,barthe2011relational,KlebanovRuemmerUlbrich2017}. At a high level, checking equivalence of two programs $P$ and $P'$ can be reduced to a sequential verification problem by executing $P'$ after $P$ and at the end checking if they always have the same outputs. It can be made more efficient by finding the right~\emph{synchronization points} and combining the code of $P$ and $P'$ in an interleaved manner. A lot of research effort in this direction is in finding good synchronization points. For the case of commutativity analysis, one can simply use the beginning of the loop as the synchronization point. However, in our experience, if we naively reduce the commutativity problem to equivalence problem and check it in a precise manner, many examples cannot be solved. In this paper, we propose a good over-approximation (c.f. Section~\ref{sec:prover}) of the reducer's behavior. The approximation allows a much more efficient yet precise enough commutativity analysis.

@@

Some reducer uses two loops to compute the output. For example, for calculating standard deviation, one first computes the average of inputs and then uses it to compute the standard deviation. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average and the other reducer assigns a non-deterministic value to the variable $a$ that is intended to store the computed average ($a := *$). The variable $a$ will be used in the computation of standard deivation. The second reducer ``over-approximates''  all possible outputs of the standard deviation computation; it computes an output even if $a$ is not the average of inputs. However, this reducer is still commutative and its commutativity verification can still be done.

@@

\textbf{The Preprocessor component} first compiles a reducer program to bytecode and use the tool Soot to further convert it to the so-called ``Jimple'' format, which is an intermediate language that is designed to simplify the analysis of Java programs. Under the Hadoop MapReduce framework, the permutation of the input is handled by the scheduler/shuffler component and is affected by things like network latency, which is not controllable by programmers. In order to deal with this issue, we wrote our own dummy Hadoop environment for the reducer as a part of the Preprocessor component so that the input order of the reducer is now controlled by J-ReCoVer. Finally, we perform in the Preprocessor some program transformation tasks to simplify the analysis. For example, we convert all output commands $out(e)$ inside loops to a special assignment (will be explained in Section~\ref{sec:program_trans1} and~\ref{sec:program_trans2}).

@@

\textbf{The BugFinder component} randomly generated 1500 pairs of lists. The lists in a pair are permutations of each other. A concrete counterexample is reported if inputting any pair of lists from the 1500 generated ones to the reducer generates different results.
Our procedure for generating random pairs is quite naive. We use five different input list of lengths 5, 7, 9, 11, and 13. For each length, we generate 300 lists and pick uniformly at random one of its permutations. Although the approach is simple, in practice it finds counterexamples in all of our non-commutative benchmarks.

@@

\subsection{The First Program Transformation Task in the Preprocessor}
\label{sec:program_trans1}
Recall that the reducer program can perform the ouput command $out(e)$ inside the loop body $s_2$. In this case, we remember the output value in a fresh variable $v$ and replace the $out(e)$ command with an assignment $v:=e$.
Notice that the value of $v$ stays the same from the point it has been assigned a value.
If the values of $v$ are the same for all permutations of input lists at the end of $s_2$, we know that the reducer is commutative w.r.t. to all output values produced inside the loop body $s_2$.

\subsection{The Prover: Reduce Commutativity Checking to SMT Solving}
\label{sec:prover}
Below we begin with a simple reducer where $s_1$ never calls the $\cur$ function. Later we will explain how to handle the general case using program transformation. The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs the values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively, for $i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way. We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$.
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

@@

In the next part, we explain how to generalize our approach to support general reducers (without the restriction on $s_1$) using program transformation.

\subsection{The Second Program Transformation Task in the Preprocessor}
\label{sec:program_trans2}

\begin{wrapfigure}{r}{3.5cm}
	\vspace{-1cm}
	\begin{algorithm}[H]
	\;\;$m := \cur + 10$; \;
	\Loop{}{
		$t:=\cur$\;
		\uIf{ $t> m$}{
			$m := t$ \;
		}
	};
	$out(m)$\;\;
\end{algorithm}
	\vspace{-0.8cm}
\caption{Reducer max$^+$}
	\vspace{-0.8cm}
	\label{fig:reducer_max1}
\end{wrapfigure}

In real-world examples, it is often that the $s_1$ part of the reducer also reads from the input. In the example in Figure~\ref{fig:reducer_max1}, the reducer program remembers the first input value in the variable $m$, increases its value by 10, and then updates its value to bigger ones if any occurs in the loop. The corresponding Equation~(\ref{eq:commu}) is valid in this case. However, the reducer is not commutative. A counterexample can be easily found. With the list $[1,2,3,4,5]$ and its permutation $[5,4,3,2,1]$ as the inputs, the reducer respectively outputs $11$ and $15$.

\subsection{Optimization: Live Variable Analysis}
\label{section:optimizations}
In this section, we explain how data flow analysis is used in J-ReCoVer to improve the precision of commutativity checking and the algorithm we used to construct the $F$ function in Equation~(\ref{eq:commu}).

@@

In order to properly evaluate the performance of J-ReCoVer, we collect benchmark example from two different sources.
First, we use the search engine \emph{searchcode.com} to collect java programs containing the key strings ``public void reduce'' or ``protected void reduce''. Since there is an upper bound on the number of results returned from the search engine, we add different search filters in order to get more data. We tried all $12$ combinations of six filters on code length $\{<50$, $ 50\sim 250$, $250\sim 450$, $450\sim 650$, $650\sim 850$, $850\sim 1050$, $1050\sim 1250\}$
and two filters for data sources $\{\url{github.com}, \url{bitbuckect.com}\}$. In total, we get $11346$ Java programs. We exclude cases that are not Hadoop MapReduce reducer programs (those do not import the Hadoop library, do not extend or implement the reducer interface) and obtain $1273$ examples. We further remove those that cannot be compiled, those that are identical to each other, or those with non-numerical data types (e.g, string) and obtained $118$ programs as the final benchmarks.

\hide{The second set of examples are those have been discussed in verification literature~\cite{ChenHSW15,ChenSW16}. This set is relatively small, we have only $10$ examples.
We also created a program that randomly generates reducer programs in order to test the scalability of J-ReCoVer.
All benchmark examples can be found on the website of J-ReCoVer.}

For the optimization based on live variable analysis (Section~\ref{section:optimizations}), we evaluate them based on real-world benchmark sets collected from open repositories (118 examples). The difference between the version with and without this optimization is very clear. The results are presented in table~\ref{tab:opt1}. The first column is the result without optimization and the second column is the result with optimization. The precision of the analysis is improved from $6.8\%$ to $97.5\%$. There are only three cases that J-ReCoVer with data flow analysis cannot handle. The reason that J-ReCoVer fails is that the three reducers use more complicated structure than what J-ReCoVer currently supports; they use a branch statement before entering the loop, i.e., in the form $\ite{g}{(s_1;\rloop\{s_2\};s_3)}{(s'_1;\rloop\{s'_2\};s'_3)}$. In theory, such a program can be handled by more sophisticated program transformation tasks. For example, we can merge the two loops and push the outer branch condition into the merged loop.
However, this new program transformation task is not implemented in J-ReCoVer yet.

@@

	\end{tabular}
	\caption{The improvement of precision using data flow analysis. The row ``Precision'' is calculated as the number of examples that J-ReCoVer returns commutative or non-commutative divided by the total number of example.}
	\label{tab:opt1}
\end{table}

>>>>

\subsection*{Related Works}
Reducer commutativity problem can be reduced to a program equivalence problem. One creates another program $R'$ that first non-deterministically swaps two consecutive input values and then executes the code of $R$. If $R'$ and $R$ are equivalent, using the fact that all permutations of a list can be obtained by swapping consecutive list elements finitely many times\footnote{Here is an example to produce [3;2;5;1;4] from [1;2;3;4;5] by swapping consecutive elements: [\textbf{1};\textbf{2};3;4;5] $\rightarrow$ [2;\textbf{1};\textbf{3};4;5] $\rightarrow$ [2;3;1;\textbf{4};\textbf{5}]$\rightarrow$ [2;3;\textbf{1};\textbf{5};4]$\rightarrow$ [\textbf{2};\textbf{3};5;1;4]$\rightarrow$ [3;2;5;1;4].}, $R$ can be proved to be commutative. There is a series of research results on program equivalence checking (or closely related topics such as regression verification and translation validation), to name a few~\cite{Pnueli:1998:TV,fedyukovich2015automated,barthe2011relational,KlebanovRuemmerUlbrich2017}. At a high level, checking equivalence of two programs $P$ and $P'$ can be reduced to a sequential verification problem by executing $P'$ after $P$ and at the end checking if they always have the same outputs. It can be made more efficient by finding the right~\emph{synchronization points} and combining the code of $P$ and $P'$ in an interleaved manner. A lot of research effort in this direction is in finding good synchronization points. For the case of commutativity analysis, one can simply use the beginning of the loop as the synchronization point. However, in our experience, if we naively reduce the commutativity problem to equivalence problem and check it in a precise manner, many examples cannot be solved. In this paper, we propose a good over-approximation (c.f. Section~\ref{sec:prover}) of the reducer's behavior. The approximation allows a much more efficient yet precise enough commutativity analysis.
Our approach uses some form of sequentialization which is, however, specialized for the reducer commutativity problem and different from the sequentialisation approaches proposed for verifying common multithreaded programs\cite{XXX}: indeed, in MapReduce programs there is no notion of threads and context switches.

@@

Some reducer uses two loops to compute the output. For example, for calculating standard deviation, one first computes the average of inputs and then uses it to compute the final result. In that case, we can still verify the reducer by first partitioning it into two reducers and then verifying them separately. One reducer computes the average and the other reducer performs the rest of the computation.
The second reducer replaces the average with a non-deterministic value and thus ``over-approximates''  all possible outputs of the standard deviation computation; it computes an output for all possible ``average'' value. However, this reducer is still commutative and its commutativity verification can still be done.

@@

\textbf{The Preprocessor component} first compiles a reducer program to bytecode and use the tool Soot to further convert it to the so-called ``Jimple'' format, which is an intermediate language that is designed to simplify the analysis of Java programs. Under the Hadoop MapReduce framework, the permutation of the input is handled by the scheduler/shuffler component and is affected by things like network latency, which is not controllable by programmers. In order to deal with this issue, we wrote our own dummy Hadoop environment for the reducer as a part of the Preprocessor component so that the input order of the reducer is now controlled by J-ReCoVer. Finally, we perform in the Preprocessor some program transformation tasks to simplify the analysis. For example, we convert all output commands $out(e)$ inside loops to a special assignment (will be explained in Section~\ref{sec:program_trans1}).

@@

\textbf{The BugFinder component} randomly generates pairs of lists. The lists in a pair are permutations of each other. A concrete counterexample is reported if inputting any pair of lists from the generated ones to the reducer generates different results.
Our procedure for generating random pairs is quite naive. We use five different input list of lengths 5, 7, 9, 11, and 13. For each length, we generate 300 lists and pick uniformly at random one of its permutations. Although the approach is simple, in practice it finds counterexamples in all of our non-commutative benchmarks in a few seconds.

@@

\subsection{The Program Transformation Task in the Preprocessor}
\label{sec:program_trans1}
\subsubsection{Handling output inside the loop body} Recall that the reducer program can perform the ouput command $out(e)$ inside the loop body $s_2$. In this case, we remember the output value in a fresh variable $v$ and replace the $out(e)$ command with an assignment $v:=e$.
Notice that the value of $v$ stays the same from the point it has been assigned a value until the end of the loop body $s_2$.
If the values of $v$ are the same for all permutations of input lists at the end of $s_2$, we know that the reducer is commutative w.r.t. to all output values produced inside the loop body $s_2$.

\begin{wrapfigure}{r}{3.5cm}
	\vspace{-1.5cm}
	\begin{algorithm}[H]
		\;\;$m := \cur + 10$; \;
		\Loop{}{
			$t:=\cur$\;
			\uIf{ $t> m$}{
				$m := t$ \;
			}
		};
		$out(m)$\;\;
	\end{algorithm}
	\vspace{-0.8cm}
	\caption{Reducer max$^+$}
	\vspace{-0.8cm}
	\label{fig:reducer_max1}
\end{wrapfigure}

\subsubsection{Handling the access to $\cur$ before entering the loop} In real-world examples, it is often that the $s_1$ part of the reducer also reads from the input. In the example in Figure~\ref{fig:reducer_max1}, the reducer program remembers the first input value in the variable $m$, increases its value by 10, and then updates its value to bigger ones if any occurs in the loop. The corresponding Equation~(\ref{eq:commu}) is valid in this case. However, the reducer is not commutative. A counterexample can be easily found. With the list $[1,2,3,4,5]$ and its permutation $[5,4,3,2,1]$ as the inputs, the reducer respectively outputs $11$ and $15$.

\subsection{The Prover: Reduce Commutativity Checking to SMT Solving}
\label{sec:prover}
After the transformation, the reducer $s_1;\rloop\{s_2\};s_3$ never calls the $\cur$ function before enthering the loop (the $s_1$ part) and never output values inside the loop body (the $s_2$ part). The command $s_2$ can be viewed as a function $F$ that reads the values of all variables and the current input before executing $s_2$ and outputs the values of all variables after $s_2$. Formally, the function $F(n,x_1,x_2,\ldots,x_k): \Z^{k+1} \rightarrow \Z^k$ returns a tuple of values $x_1',x_2',\ldots,x_k'$, where $n$ is the current input value of the reducer, $x_i$ and $x_i'$ are the values of variables before and after the execution of $s_2$, respectively, for $i\in[1,k]$. The construction of $F$ from $s_2$ can be done in the standard way. We check if the formula below is valid for all possible values of $n_1,n_2, x_1,x_2,\ldots,x_k$.
\begin{equation}
 F(n_1, F(n_2,x_1,x_2,\ldots,x_k)) = F(n_2, F(n_1,x_1,x_2,\ldots,x_k) )
\label{eq:commu}
\end{equation}

@@

\subsection{Optimization: Live Variable Analysis}
\label{section:optimizations}
Here we explain how live variable analysis is used in J-ReCoVer to improve the precision of commutativity checking.

@@

In order to properly evaluate the performance of J-ReCoVer, We use the search engine \emph{searchcode.com} to collect java programs containing the key strings ``public void reduce'' or ``protected void reduce''. Since there is an upper bound on the number of results returned from the search engine, we add different search filters in order to get more data. We tried all $12$ combinations of six filters on code length $\{<50$, $ 50\sim 250$, $250\sim 450$, $450\sim 650$, $650\sim 850$, $850\sim 1050$, $1050\sim 1250\}$
and two filters for data sources $\{\url{github.com}, \url{bitbuckect.com}\}$. In total, we get $11346$ Java programs. We exclude cases that are not Hadoop MapReduce reducer programs (those do not import the Hadoop library, do not extend or implement the reducer interface) and obtain $1273$ examples. We further remove those that cannot be compiled, those that are identical to each other, or those with non-numerical data types (e.g, string) and obtained $118$ programs as the final benchmarks.

There are only three cases that J-ReCoVer with live variable analysis cannot handle. The reason that J-ReCoVer fails is that the three reducers use more complicated structure than what J-ReCoVer currently supports; they use a branch statement before entering the loop, i.e., in the form $\ite{g}{(s_1;\rloop\{s_2\};s_3)}{(s'_1;\rloop\{s'_2\};s'_3)}$. In theory, such a program can be handled by more sophisticated program transformation tasks. For example, we can merge the two loops and push the outer branch condition into the merged loop.

J-ReCoVer is the first tool specialized to the reducer commutativity problem, although the approach is simple, it solves most of the real benchmarks we collected from repositories
There is no other tool that could handle the reducers as they are. Perhaps some other tools could be applied on the transformed programs, but the transformations would still be needed, and our SMT-based backend verifier turned out to work efficiently, and so we did not feel a need to replace it by another verification tool. Of course, in the future, this can be done if need be.

\hide{
For the optimization based on live variable analysis (Section~\ref{section:optimizations}), we evaluate them based on real-world benchmark sets collected from open repositories (118 examples). The difference between the version with and without this optimization is very clear. The results are presented in table~\ref{tab:opt1}. The first column is the result without optimization and the second column is the result with optimization. The precision of the analysis is improved from $6.8\%$ to $97.5\%$. 
However, this new program transformation task is not implemented in J-ReCoVer yet.

@@

	\end{tabular}
	\caption{The improvement of precision using data flow analysis. The row ``Precision'' is calculated as the number of examples that J-ReCoVer returns commutative or non-commutative divided by the total number of example.}
	\label{tab:opt1}
\end{table}
}
